### 2021-05-29 papers/thesis/full/ch4-spherical-caps/spherical-caps.tex
Further edits. Included more references.

The spherical cap can be thought of as higher dimensional version of the circle arc (a one-dimensional \enquote{surface} in two-dimensional space). One can define OPs on the circle in terms of Fourier series, which we write here as orthogonal polynomials in $x$ and $y$:
\begin{definition}\label{def:sc:Ydefinition}
	Define the unit circle $\omega := \{ \mathbf{x} = (x,y) \in \R^2 \: | \: x^2 + y^2 = 1\}$, and define the parameter $\theta$ for each $(x,y) \in \omega$ by $x = \cos\theta$, $y = \sin\theta$. Define the polynomials $\{\chki\}$ for $k = 0, 1, \dots$, $i = 0, 1$ on $(x,y) \in \omega$ by
\bseqn{
	\ch_{0,0}(\xvec) \equiv \ch_{0,0}(x, y) &:= \ch_{0} =: \ch_{0,0}(\theta) \\
	\ch_{k,0}(\xvec) \equiv \ch_{k,0}(x, y) \ &:= T_k(x) = \cos k \theta =: \ch_{k,0}(\theta), \quad k = 1,2,3,\dots \\
	\ch_{k,1}(\xvec) \equiv \ch_{k,1}(x, y) &:= y \: U_{k-1}(x) = \sin k \theta =: \ch_{k,1}(\theta), \quad k = 1,2,3,\dots
}
where $\ch_{0} := \frac{\sqrt{2}}{2}$ and $T_k$, $U_{k-1}$ are the standard Chebyshev polynomials on the interval $[-1,1]$. The $\{Y_{k,i}\}$ are orthonormal with respect to the inner product
\bseqn{
	\ip< p, \: q >_{\ch} &:= \frac{1}{\pi} \: \int_0^{2\pi} p(\xvec(\theta)) \: q(\xvec(\theta)) \: \D \theta.
}
Note that $\ch_0$ is defined so as to ensure orthonormality. 
\end{definition}

However, for an arc of the unit circle $\{ define arc \}$ (note we can orientate the arc as we please without loss of generality), we would need to make a modification so that the polynomials are orthogonal with respect to an inner product integral with the truncated domain of the arc:
\bseqn{
	give arc ops (T^h, yU^h) and the inner product.
}
\bstodoinline{Expand on this. Read the Olver Xu "OPs on quadratic curves paper." where the arc might apply??? Either way, cite this paper in intro (CH1).}

@@

This idea of using 1D OPs on a shortened interval to create the modified OPs for an arc has be generalised to higher dimensional domains, where a way of constructing orthogonal polynomials on algebraic curves has been presented \cite{olver2020orthogonal}. \bstodoinline{Expand on this. Summarise the "on" surfaces approach. Use references in the paper.}

@@

Using the approach presented for constructing OPs on algebraic curves, we can construct the 3D orthogonal polynomials on the spherical cap or band $\Omega$ from 1D orthogonal polynomials on the interval \([\alpha,\beta]\), and from Fourier series written as orthogonal polynomials in $x$ and $y$:
\begin{proposition}[\cite{olver2020orthogonal}]\label{prop:sc:construction}
	Let $w : (\alpha,\beta) \to \R$ be a weight function. For $n = 0,1,2,\dots, $ let $\{r_{n,k}\}$ be polynomials orthogonal with respect to the weight $\rho(x)^{2k} w(x)$ where $0 \le k \le n$. Then the 3D polynomials defined on $\Omega$ given by
\bseqn{
	\scopnki(x,y,z) := r_{n-k,k}(z) \: \rho(z)^k \: \ch_{k,i}\fpr({x \over \rho(z)}, {y \over \rho(z)})
}
for $i \in {0,1}, \: 0 \le k \le n, \: n = 0,1,2,\dots$ are orthogonal polynomials with respect to the inner product
\bseqn{
	\ip< p, \: q > &:= \int_\Omega p(x,y,z) \: q(x,y,z) \: w(z) \: \D A \\
	&= \int_0^{\cos^{-1}(\alpha)} \int_0^{2\pi} p\big(\sinphi \costheta, \sinphi \sintheta, \cosphi \big) \: q\big(\sinphi \costheta, \sinphi \sintheta, \cosphi \big) \: \times \\
	&\qquad w(\cosphi) \: \sinphi \: \D \theta \: \D \varphi \\
	&= \int_\alpha^1 \int_0^{2\pi} p\big(\rho(z) \costheta, \rho(z) \sintheta, z\big) \: q\big(\rho(z) \costheta, \rho(z) \sintheta, z\big) \: w(z) \: \D \theta \: \D z
}
on $\Omega$, where $\D A = \sinphi \: \D \theta \: \D \varphi$ is the uniform spherical measure on $\Omega$. 
\end{proposition}

>>>>

The spherical cap can be thought of as higher dimensional version of the circle arc (a one-dimensional \enquote{surface} in two-dimensional space). One can define OPs on the circle in terms of Fourier series, which we write here as orthogonal polynomials in $x$ and $y$:
\begin{definition}\label{def:sc:Ydefinition}
	Define the unit circle $\omega := \bsset{\vec{x} = (x,y) \in \R^2}{x^2 + y^2 = 1}$, and define the parameter $\theta$ for each $(x,y) \in \omega$ by $x = \cos\theta$, $y = \sin\theta$. Define the polynomials $\{\chki\}$ for $k = 0, 1, \dots$, $i = 0, 1$ on $(x,y) \in \omega$ by
\bseqn{
	\ch_{0,0}(\xvec) \equiv \ch_{0,0}(x, y) &:= \ch_{0} =: \ch_{0,0}(\theta) \\
	\ch_{k,0}(\xvec) \equiv \ch_{k,0}(x, y) &:= T_k(x) = \cos k \theta =: \ch_{k,0}(\theta) \\
	\ch_{k,1}(\xvec) \equiv \ch_{k,1}(x, y) &:= y \: U_{k-1}(x) = \sin k \theta =: \ch_{k,1}(\theta)
}
for $k = 1,2,3,\dots$ where $\ch_{0} := {\sqrt{2} \over 2}$ and $T_k$, $U_{k-1}$ are the standard Chebyshev polynomials on the interval $[-1,1]$. The $\{Y_{k,i}\}$ are orthonormal with respect to the inner product
\bseqn{
	\ip< p, \: q >_{\ch} &:= \frac{1}{\pi} \: \int_0^{2\pi} p(\xvec(\theta)) \: q(\xvec(\theta)) \: \D \theta.
}
Note that $\ch_0$ is defined so as to ensure orthonormality. 
\end{definition}

However, for an arc of the unit circle $\bsset{\vec{x} = (x,y) \in \R^2}{x \ge h, \: x^2 + y^2 = 1}$ for some $h \in (-1, 1)$ (note we can orientate the arc as we please without loss of generality), we would need to make a modification so that the polynomials are orthonormal with respect to a different inner product accounting for the truncated domain of the arc:
\bseqn{
	\ip< p, \: q >_{\ch^h} &:= c^h \: \int_{-\arccos{h}}^{\arccos{h}} p(\xvec(\theta)) \: q(\xvec(\theta)) \: \D \theta \\
	&= c^h \: \int_h^1 \Big\{ p(x,\sqrt{1-x^2}) \; q(x,\sqrt{1-x^2}) \\
	&\quadfour \quadtwo + p(x,-\sqrt{1-x^2}) \; q(x,-\sqrt{1-x^2}) \Big\} {\D x \over \sqrt{1-x^2}}
}
for some normalising constant $c^h$. In the same vein, we can choose to construct the arc OPs in the same way as the whole circle OPs -- that is, define $\{T_k^h\}$, $\{U_k^h\}$ as two sets of univariate orthogonal polynomials on the interval $[h,1]$, orthogonal with respect to the weight functions $(1-x^2)^{-\half}$ and $(1-x^2)^{\half}$ respectively, and then define
\bseqn{
	\ch^h_{0,0}(\xvec) \equiv \ch^h_{0,0}(x, y) &:= \ch^h_{0} =: \ch^h_{0,0}(\theta) \\
	\ch^h_{k,0}(\xvec) \equiv \ch^h_{k,0}(x, y) &:= T^h_k(x) =: \ch^h_{k,0}(\theta) \\
	\ch^h_{k,1}(\xvec) \equiv \ch^h_{k,1}(x, y) &:= y \: U^h_{k-1}(x) =: \ch^h_{k,1}(\theta)
}
where $Y_0^h$ is chosen so that $\ip< Y_0^h, \: Y_0^h >_{\ch^h} = 1$.

\bstodoinline{Expand on this. Read the Olver Xu "OPs on quadratic curves paper." where the arc might apply??? Either way, cite this paper in intro (CH1).}

@@

This idea of using 1D OPs on a shortened interval to create the modified OPs for a circle arc has be generalised to higher dimensional domains, where a way of constructing orthogonal polynomials on algebraic curves has been presented \cite{olver2020orthogonal}. \bstodoinline{Expand on this. Summarise the "on" surfaces approach. Use references in the paper.}

@@

Using the approach presented for constructing OPs on algebraic curves, we can construct the 3D orthogonal polynomials on the spherical cap or band $\Omega$ from 1D orthogonal polynomials on the interval $[\alpha,\beta]$, and from Fourier series written as orthogonal polynomials in $x$ and $y$. Recall that the spherical cap is defined by 
\bseqn{
	\Omega := \bsset{(x,y,z) \in \R^3}{\alpha < z < \beta, \: x^2 + y^2 + z^2 = 1}.
}
\begin{proposition}[\cite{olver2020orthogonal}]\label{prop:sc:construction}
	Let $w : (\alpha,\beta) \to \R$ be a weight function. For $n = 0,1,2,\dots, $ let $\{r_{n,k}\}$ be polynomials orthogonal with respect to the weight $\rho(x)^{2k} w(x)$ where $0 \le k \le n$. Then the 3D polynomials defined on $\Omega$ given by
\bseqn{
	\scopnki(x,y,z) := r_{n-k,k}(z) \: \rho(z)^k \: \ch_{k,i}\fpr({x \over \rho(z)}, {y \over \rho(z)})
}
for $i \in {0,1}, \: 0 \le k \le n, \: n = 0,1,2,\dots$ are orthogonal polynomials with respect to the inner product
\bseqn{
	&\ip< p, \: q > := \int_\Omega p(x,y,z) \: q(x,y,z) \: w(z) \: \D A \\
	&= \int_0^{\cos^{-1}(\alpha)} \int_0^{2\pi} \Big\{ p\big(\sinphi \costheta, \sinphi \sintheta, \cosphi \big) \: q\big(\sinphi \costheta, \sinphi \sintheta, \cosphi \big) \\
	&\quadeight \quad \cdot w(\cosphi) \: \sinphi \Big\} \: \D \theta \: \D \varphi \\
	&= \int_\alpha^1 \int_0^{2\pi} p\big(\rho(z) \costheta, \rho(z) \sintheta, z\big) \: q\big(\rho(z) \costheta, \rho(z) \sintheta, z\big) \: w(z) \: \D \theta \: \D z
}
on $\Omega$, where $\D A = \sinphi \: \D \theta \: \D \varphi$ is the uniform spherical measure on $\Omega$. 
\end{proposition}
